\noindent UNIX is a trademark of X/Open\\
MS-DOS and Microsoft Windows are trademarks of Microsoft Corporation\\
OS/2 and Operating System/2 are trademarks of IBM\\
X Window System is a trademark of X Consortium, Inc.\\
Motif is a trademark of the Open Software Foundation\\
\linux\ is not a trademark, and has no connection to
UNIX, Unix System Labratories, or to X/Open.\\
Please bring all unacknowledged trademarks to the attention of the author.
\vspace*{\fill}

\noindent
Copyright \copyright \ \years \ Larry Greenfield\\
427 Harrison Avenue\\Highland Park, NJ\\08904\\
{\tt leg+@andrew.cmu.edu}

\vspace{.2in}\noindent
Permission is granted to make and distribute verbatim copes of this
manual provided the copyright notice and this permission notice are
preserved on all copies.

\noindent
Permission is granted to copy and distribute modified versions of this
manual under the conditions for verbatim copying, provided also that
the sections that reprint ``The GNU General Public License'', ``The
GNU Library General Public License'', and other clearly marked
sections held under seperate copyright are reproduced under the
conditions given within them, and provided that the entire resulting
derived work is distributed under the terms of a permission notice
identical to this one.

\noindent 
Permission is granted to copy and distribute translations of
this manual into another language under the conditions for modified
versions.  ``The GNU General Public License'' and ``The GNU Library
General Public License'' may be included in a translation approved by
the Free Software Foundation instead of in the original English.

\noindent
At your option, you may distribute verbatim and modified versions of
this document under the terms of the GNU General Public License,
excepting the clearly marked sections held under seperate copyright.

Exceptions to these rules may be granted for various purposes: Write
to Larry Greenfield at the above address or email {\tt
  leg+@andrew.cmu.edu}, and ask.  It is requested (but not
required) that you notify the author whenever commercially or
large-scale printing this document.  Royalties and donations are
accepted and will encourage further editions.

